---
title: "Hotel Booking Demand Anlysis"
author: "Hongyu Wang (Olivia)"
date: "12/22/2020"
output: github_document
---

### PART1 Introduction
In my project, I am working with hotel booking demand available on Kaggle, which contains booking information for a city hotel and a resort hotel, and includes information such as when the booking was made, arrival date, length of stay, the number of adults, children, and/or babies, and the number of available parking spaces, among other things.

The varibales and the description of the values that I included from the original data are as follows:

1. hotel: Hotel (H1 = Resort Hotel or H2 = City Hotel)
2. is_canceled: Value indicating if the booking was canceled (1) or not (0)
3. lead_time: Number of days that elapsed between the entering date of the booking into the PMS and the arrival date
4. arrival_date_year: Year of arrival date
5. arrival_date_month: Month of arrival date
6. arrival_date_day_of_month: Day of arrival date
7. stays_in_weekend_nights: Number of weekend nights (Saturday or Sunday) the guest stayed or booked to stay at the hotel
8. stays_in_week_nights: Number of week nights (Monday to Friday) the guest stayed or booked to stay at the hotel
9. adults: Number of adults
10. children: Number of children
11. babies: Number of babies
12. mealType of meal booked. Categories are presented in standard hospitality meal packages: Undefined/SC – no meal package; BB – Bed & Breakfast; HB – Half board (breakfast and one other meal – usually dinner); FB – Full board (breakfast, lunch and dinner)
13. country: Country of origin. Categories are represented in the ISO 3155–3:2013 format
14. distribution_channel: Booking distribution channel. The term “TA” means “Travel Agents” and “TO” means “Tour Operators”
15. is_repeated_guestValue indicating if the booking name was from a repeated guest (1) or not (0)
16. previous_cancellations: Number of previous bookings that were cancelled by the customer prior to the current booking
17. previous_bookings_not_canceled: Number of previous bookings not cancelled by the customer prior to the current booking
18. reserved_room_type: Code of room type reserved. Code is presented instead of designation for anonymity reasons.
19. required_car_parking_spaces: Number of car parking spaces required by the customer


### PART2 Preparing the data
* Load required packages and import data
```{r setup}
# Set up R markdown
knitr::opts_chunk$set(echo = TRUE, fig.align='center', warning = FALSE, 
                      message = FALSE)
# Load in required packages
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
suppressMessages(library(GGally))
suppressMessages(library(gridExtra))
suppressMessages(library(sampling))
# read dataset
df <- read.csv("/Users/hongyuwang/Documents/R_project/Hotel_booking/hotel_bookings.csv")
```

* structure of the dataframe
```{r}
str(df)
```

From the structure of the dataframe, I think combining arrival_date_year, arrival_date_month and arrival_date_day_of_month these three columns into one single coulmn - arrival_date would be necessary for my analysis. 

```{r}
df$arrival_date_month <- ordered(df$arrival_date_month, 
                                 levels = c("January", "February", "March", "April", "May",
                                            "June", "July", "August", "September", "October",
                                            "November", "December"))
df <- df %>% mutate(arrival_date = make_date(arrival_date_year, 
                                             arrival_date_month,  
                                             arrival_date_day_of_month))
head(df$arrival_date)
```

* Dealing with NAs
```{r}
# extract columns that we need
df <- df[c("hotel", "is_canceled", "lead_time", "arrival_date", 
                                   "arrival_date_month", "arrival_date_year",
                                   "stays_in_weekend_nights", "stays_in_week_nights",
                                   "adults", "children", "babies", "meal", "country",
                                   "distribution_channel", "is_repeated_guest",
                                   "previous_cancellations",
                                   "previous_bookings_not_canceled", "reserved_room_type",
                                   "required_car_parking_spaces")]
# check whether the data contains any NAs
any(is.na(df))
```
The result shows that there are any NAs in our data frame and therefore we need to get rid of them. Also, I only keep the columns that we are interested in.
```{r results='hide'}
# remove rows containing NAs
hotel_bookings <- df[complete.cases(df),]
any(is.na(hotel_bookings))
```
the output - FALSE indicates that we've removed NAs successfully.

* summary of dataframe
```{r}
# Look at the summary of statistics for the dataframe
summary(hotel_bookings)
```

From the summay of hotel_bookings. we can see that there is no obvious data error such as negative data.  
This data summary reveals some useful information:  
1. 37.04% of booking are cancled according to the mean value for the is_canceld column.  
2. People who booking the hotel are mostly adults and very small amount of children and babies.  
3. The most populat type of meal is Bed & Breakfast, while the least popular one is Full Board(breakfast, lunch and dinner). Among all the type of meals, only FB contains lunch, so hotels may prepare less for the lunch.  
4. Most of the customers who book the hotel are new customers.  
5. Most customers do not need a car parking space,   
  


### PART3 Analyzing the Data
#### Question 1:  whether there are any strong correlations between certain varibales, 
Assumption: maybe when there are children, people are incline to go to the hotel on weekends? Maybe people who have canclelation history record are more likely to cancel the booking?
```{r}
ggcorr(as.matrix(hotel_bookings[, c("is_canceled", "stays_in_weekend_nights",
                                    "stays_in_week_nights", "adults", "children",
                                    "babies", "is_repeated_guest", "previous_cancellations",
                                    "previous_bookings_not_canceled")]), label = TRUE, hjust = 0.8)

```

From the correlation result we've got, there are no obvious correlation (mostly less than 0.1) between these variables. The 0.5 correaltion between stay_in_week_nights and stays_in_weekend_nights may indicates that people may incline to book hotel from week days to weekends.<br/><br/>


#### Question 2: Which are the busiest months for hotels?
```{r}
### year 2015
booking_by_month <- hotel_bookings %>%
  subset(arrival_date_year == 2015) %>%
  group_by(arrival_date_month) %>%
  summarize(n = n())
# design labs for the plot
month_labels = c('Jul' ,'Aug', 'Sep', 'Oct', 'Nov' , 'Dec')
booking_by_month$month_name <- month_labels
booking_by_month$month_name <- factor(booking_by_month$month_name,
                                      levels = unique(booking_by_month$month_name))
p1 <- ggplot(booking_by_month, aes(x = month_name, y = n)) + 
  geom_bar(stat = 'identity', fill="navyblue", color = 'orange', width = 1, lwd = 0.8) +
  labs(x = 'Arrival Month', y = "Total Customers",
       title = 'Book Amount by Month of Year 2015') +
  geom_point(shape=21, fill="red", color = 'black') +
  geom_line(group = 1, size = 1, color = 'orange') +
  theme_bw() + 
  theme(plot.title = element_text(hjust=0.5))

### year 2016
booking_by_month <- hotel_bookings %>%
  subset(arrival_date_year == 2016) %>%
  group_by(arrival_date_month) %>%
  summarize(n = n())
# design labs for the plot
month_labels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul' ,'Aug', 'Sep', 
                 'Oct', 'Nov' , 'Dec')
booking_by_month$month_name <- month_labels
booking_by_month$month_name <- factor(booking_by_month$month_name,
                                      levels = unique(booking_by_month$month_name))
p2 <- ggplot(booking_by_month, aes(x = month_name, y = n)) + 
  geom_bar(stat = 'identity', fill="navyblue", color = 'orange', width = 1, lwd = 0.8) +
  labs(x = 'Arrival Month', y = "Total Customers",
       title = 'Book Amount by Month of Year 2016') +
  geom_point(shape=21, fill="red", color = 'black') +
  geom_line(group = 1, size = 1, color = 'orange') +
  theme_bw() + 
  theme(plot.title = element_text(hjust=0.5))

### 2017
booking_by_month <- hotel_bookings %>%
  subset(arrival_date_year == 2017) %>%
  group_by(arrival_date_month) %>%
  summarize(n = n())
# design labs for the plot
month_labels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul' ,'Aug')
booking_by_month$month_name <- month_labels
booking_by_month$month_name <- factor(booking_by_month$month_name,
                                      levels = unique(booking_by_month$month_name))
p3 <- ggplot(booking_by_month, aes(x = month_name, y = n)) + 
  geom_bar(stat = 'identity', fill="navyblue", color = 'orange', width = 1, lwd = 0.8) +
  labs(x = 'Arrival Month', y = "Total Customers",
       title = 'Book Amount by Month of Year 2017') +
  geom_point(shape=21, fill="red", color = 'black') +
  geom_line(group = 1, size = 1, color = 'orange') +
  theme_bw() + 
  theme(plot.title = element_text(hjust=0.5))
# combine the three plots
grid.arrange(p1, p2, p3, ncol = 1)
```
We can conclude that the busiest months are from March to June, and from September to October.<br /> <br /> 
 

#### Qestion 3: Does cancel rate has similar distribution as booking amount does?

```{r}
cancel_by_month <- hotel_bookings %>%
                   subset(arrival_date_year == 2015) %>%
          # group the dataframe by arrive_date and calculate the cancle rate of each day 
                   group_by(arrival_date_month) %>%
                   summarize(n = n()) %>%
          # add field with cancel rate        
                   mutate(cancel_percent = n / nrow(hotel_bookings) * 100)

month_labels = c('Jul' ,'Aug', 'Sep', 
                 'Oct', 'Nov' , 'Dec')
cancel_by_month$month_name <- month_labels
cancel_by_month$month_name <- factor(cancel_by_month$month_name,
                                    levels = unique(cancel_by_month$month_name))
# visualize by bar plot
p1 <- ggplot(cancel_by_month, aes(x = month_name, y = cancel_percent)) +
        geom_bar(stat = 'identity', fill="orange", color = 'navyblue', 
                 width = 1, lwd = 0.8) +
        geom_point(shape=21, fill="hotpink1", color = 'black') +
        geom_line(group = 1, size = 1, color = 'steelblue4') +
        theme_bw() +
        labs(x = 'Date', y = '% of Cancel', title = 'Cancel Rate by Month of Year 2015') +
        theme(plot.title = element_text(hjust=0.5))

cancel_by_month <- hotel_bookings %>%
                   subset(arrival_date_year == 2016) %>%
          # group the dataframe by arrive_date and calculate the cancle rate of each day 
                   group_by(arrival_date_month) %>%
                   summarize(n = n()) %>%
          # add field with cancel rate        
                   mutate(cancel_percent = n / nrow(hotel_bookings) * 100)

month_labels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul' ,'Aug', 'Sep', 
                 'Oct', 'Nov' , 'Dec')
cancel_by_month$month_name <- month_labels
cancel_by_month$month_name <- factor(cancel_by_month$month_name,
                                    levels = unique(cancel_by_month$month_name))
# visualize by bar plot
p2 <- ggplot(cancel_by_month, aes(x = month_name, y = cancel_percent)) +
        geom_bar(stat = 'identity', fill="orange", color = 'navyblue', 
                 width = 1, lwd = 0.8) +
        geom_point(shape=21, fill="hotpink1", color = 'black') +
        geom_line(group = 1, size = 1, color = 'steelblue4') +
        theme_bw() +
        labs(x = 'Date', y = '% of Cancel', title = 'Cancel Rate by Month of Year 2016') +
        theme(plot.title = element_text(hjust=0.5))

cancel_by_month <- hotel_bookings %>%
                   subset(arrival_date_year == 2017) %>%
          # group the dataframe by arrive_date and calculate the cancle rate of each day 
                   group_by(arrival_date_month) %>%
                   summarize(n = n()) %>%
          # add field with cancel rate        
                   mutate(cancel_percent = n / nrow(hotel_bookings) * 100)

month_labels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul' ,'Aug')
cancel_by_month$month_name <- month_labels
cancel_by_month$month_name <- factor(cancel_by_month$month_name,
                                    levels = unique(cancel_by_month$month_name))
# visualize by bar plot
p3 <- ggplot(cancel_by_month, aes(x = month_name, y = cancel_percent)) +
        geom_bar(stat = 'identity', fill="orange", color = 'navyblue', 
                 width = 1, lwd = 0.8) +
        geom_point(shape=21, fill="hotpink1", color = 'black') +
        geom_line(group = 1, size = 1, color = 'steelblue4') +
        theme_bw() +
        labs(x = 'Date', y = '% of Cancel', title = 'Cancel Rate by Month of Year 2017') +
        theme(plot.title = element_text(hjust=0.5))

grid.arrange(p1, p2, p3, ncol = 1)
```

Compare the cancel rate graph with the booking amount graph, we can see that the plot trend are pretty much the same, and so we coonclude that book amount and cancel rate have positive relationship.<br /> <br /> 


#### Question 4: What is the distribution of number of customers by day?
```{r}
# plot for year 2015
p1 <- hotel_bookings %>% 
        subset(arrival_date_year == 2015) %>%
        group_by(arrival_date, hotel) %>%
        summarize(n = n()) %>%
        ggplot(aes(x = arrival_date, y = n)) +
        geom_point(aes(color = hotel)) +
        scale_color_manual(name='Hotel Type', values =c('navyblue','orange')) +
        labs(x = 'Arrival Date', y = 'Total Customers', title = 'Arrival Date vs. Booking Amount of Year 2015') +
        scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5))

# plot for year 2016
p2 <- hotel_bookings %>% 
        subset(arrival_date_year == 2016) %>%
        group_by(arrival_date, hotel) %>%
        summarize(n = n()) %>%
        ggplot(aes(x = arrival_date, y = n)) +
        geom_point(aes(color = hotel)) +
        scale_color_manual(name='Hotel Type', values =c('navyblue','orange')) +
        labs(x = 'Arrival Date', y = 'Total Customers', title = 'Arrival Date vs. Booking Amount of Year 2016') +
        scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5))

# plot for year 2017
p3 <- hotel_bookings %>% 
        subset(arrival_date_year == 2017) %>%
        group_by(arrival_date, hotel) %>%
        summarize(n = n()) %>%
        ggplot(aes(x = arrival_date, y = n)) +
        geom_point(aes(color = hotel)) +
        scale_color_manual(name='Hotel Type', values =c('navyblue','orange')) +
        labs(x = 'Arrival Date', y = 'Total Customers', title = 'Arrival Date vs. Booking Amount of Year 2017') +
        scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5))

# combine three plots
grid.arrange(p1, p2, p3, ncol = 1)
```

From the graph, we can see that:  
1. City hotel is more popular than resort hotel.  
2. Hotel attract more customers with year goes by.  
3. There are several days for which the total customers is close to 300, which may due to holodays.<br/><br/>  

#### Question 5: Will holiday boost the number of customers?
```{r}
# year 2015
p1 <- hotel_bookings %>% 
  subset(arrival_date_year == 2015) %>%
  group_by(arrival_date, hotel) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = arrival_date, y = n)) + 
  geom_point(aes(color = hotel)) + 
  scale_color_manual(name='Hotel Type', values =c('navyblue','orange')) +
  labs(x = 'Arrival Date', y = 'Booking Amount', title = 'Arraval Date vs. Booking Amount of Year 2015') +
  scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2015-12-25")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2015-12-25")), y= 100, label='Christmas'),
            size=4.5, angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2015-07-03")))), 
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x=(as.Date("2015-07-03")), y= 100, label='Independence Day'),
            size=4.5, angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2015-11-26")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2015-11-26")), y= 100, label='Thanksgiving'),
            size=4.5, angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2015-10-31")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2015-10-31")), y= 100, label='Halloween'),
            size=4.5, angle=90, vjust=-0.4, hjust=0)
  
# year2016
p2 <- hotel_bookings %>% 
  subset(arrival_date_year == 2016) %>%
  group_by(arrival_date, hotel) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = arrival_date, y = n)) +
  geom_point(aes(color = hotel)) +
  scale_color_manual(name='Hotel Type', values =c('navyblue','orange')) +
  labs(x = 'Arrival Date', y = 'Booking Amount', title = 'Arraval Date vs. Booking Amount of Year 2016') +
  scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-12-25")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2016-12-25")), y= 100, label='Christmas'),
            angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-01-01")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2016-01-01"), y = 100, label = 'New Year\'s Day',
            angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-01-20")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2016-01-20"), y = 100, label = 'Martin Luther King Day',
            angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-02-14")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2016-02-14"), y = 100, label = 'Valentine\'s Day',
            angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-03-17")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2016-03-17"), y = 100, label = 'St.Patrick\'s Day',
            angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-05-01")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2016-05-01")), y= 100, label='May Day'),
            angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-07-03")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2016-07-03")), y= 100, label='Independence Day'),
            angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-11-26")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2016-11-26")), y= 100, label='Thanksgiving'),
            angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2016-10-31")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2016-10-31")), y= 100, label='Halloween'),
            angle=90, vjust=-0.4, hjust=0)

# year 2017
p3 <- hotel_bookings %>% 
  subset(arrival_date_year == 2017) %>%
  group_by(arrival_date, hotel) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = arrival_date, y = n)) +
  geom_point(aes(color = hotel)) +
  scale_color_manual(name='Hotel Type', values =c('navyblue','orange')) +
  labs(x = 'Arrival Date', y = 'Booking Amount', title = 'Arraval Date vs. Number of Customers of Year 2017') +
  scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_vline(aes(xintercept=(as.numeric(as.Date("2017-01-01")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2017-01-01"), y = 100, label = 'New Year\'s Day',
                size=4.5,angle = 90, vjust=-0.4, hjust=0)) + 
  geom_vline(aes(xintercept=(as.numeric(as.Date("2017-01-20")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2017-01-20"), y = 100, label = 'Martin Luther King Day',
                size=4.5,angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2017-02-14")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2017-02-14"), y = 100, label = 'Valentine\'s Day',
                size=4.5,angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2017-03-17")))),
             linetype = 2, color = 'brown', lwd = 0.8) +
  geom_text(aes(x = as.Date("2017-03-17"), y = 100, label = 'St.Patrick\'s Day',
                size=4.5,angle = 90, vjust=-0.4, hjust=0)) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2017-05-01")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2017-05-01")), y= 100, label='May Day'),
                size=4.5,angle=90, vjust=-0.4, hjust=0) +
  geom_vline(aes(xintercept=(as.numeric(as.Date("2017-07-03")))), 
             linetype = 2, color = 'brown', lwd = 0.8) + 
  geom_text(aes(x=(as.Date("2017-07-03")), y= 100, label='Independence Day'),
                size=4.5,angle=90, vjust=-0.4, hjust=0)

p1
p2
p3

```

In year 2015, there is no obvious correlation between holidays and the number of customers, while in year 2016 and year 2017, holidays appear to line up with high number of customers.<br/><br/>  

#### Qestion 6: Does longer lead time lead to higher cancel rate?  
```{r}
# filter all the outliers of lead time
f <- fivenum(hotel_bookings$lead_time)
outlier <- c(f[2] - 1.5*(f[4] - f[2]), f[4] + 1.5*(f[4] - f[2]))
rows <- which(hotel_bookings$lead_time >= outlier[1] & hotel_bookings$lead_time <= outlier[2])
lead_time_vs_cancel <- hotel_bookings[rows, c("lead_time", "is_canceled")]
lead_time_vs_cancel <- lead_time_vs_cancel %>%
                       group_by(lead_time) %>%
                       summarize(cancel_rate = mean(is_canceled),
                                 n = n())

ggplot(lead_time_vs_cancel, aes(x = lead_time, y = cancel_rate)) + 
  geom_point(color = 'navyblue') +
  geom_smooth(method = 'lm', color = 'orange', se = FALSE)+
  labs(x = "Lead Time", y = "Cancle Rate", title = "Lead Time vs. Cancel Rate") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

# cut the data into bind of 20days
range(lead_time_vs_cancel$lead_time)
lead_time_vs_cancel$bucket <- cut(lead_time_vs_cancel$lead_time,
                                  breaks = c(0, 5, 10, 30, 50, 100, 200, 300, 400), inclued.lowest = TRUE, right = FALSE)
# create a new dataframe grouping by lead time bucket
lead_time_vs_cancel_bucket <- lead_time_vs_cancel %>%
                              group_by(bucket) %>%
                              summarize(cancel_rate = mean(cancel_rate),
                                        n = n())
lead_time_vs_cancel_bucket$relative_cancel_rate <- 
  with(lead_time_vs_cancel_bucket, 100 * (cancel_rate - mean(cancel_rate)) / mean(cancel_rate))
# visualization
ggplot(lead_time_vs_cancel_bucket, aes(x = bucket, y = relative_cancel_rate)) +
  geom_bar(stat = 'identity', color = "darkorange3", fill = "royalblue4", width = 1, lwd = 0.8) +
  labs(x = "Lead Time Group (days)", y = "Relative Cancel Rate %",
       title = "Cancel Rate vs Lead Time Group") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
cor.test(lead_time_vs_cancel$lead_time, lead_time_vs_cancel$cancel_rate)

```

From the charts, the longer the lead time is, the higher cancel rate appears. Also, when lead time is larger than 50 days, cancel rate will be above average cancel rate. Hotels may need to .add extra number of available booking rooms to offset the reduced profit due to canceled bookings. <br/><br/>


#### Qestion 7 : How long do customers usually stay in the hotel?
```{r}
total_nights <- hotel_bookings$stays_in_weekend_nights + hotel_bookings$stays_in_week_nights
total_nights <- as.data.frame(table(total_nights))
total_nights <- arrange(total_nights, desc(Freq))
ggplot(total_nights[1:10, ], aes(x = total_nights, y = Freq,
                                 fill = ifelse(total_nights == 2, "Highlight", "Normal"))) + 
  geom_bar(stat = 'identity', color = "white", width = 1, lwd = 0.8, show.legend = FALSE) +
   scale_fill_manual(values =c('orange', 'navyblue')) +
  labs(x = "Total Nights", y = "Frequency", title = 'Frequency of Total Stay Night (Top 10)') +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

Based on the chart, most customers stay in the hotel for 1 to 3 days, and 2 day is the most popular one.<br/><br/>

#### Question 8: Examine the distribution of the lead time

```{r}
ggplot(hotel_bookings, aes(x = lead_time)) + 
  geom_histogram(color = "orange", fill = "navyblue") +
  theme_bw() +
  labs(x = "Lead Time (days)", y = "Frequency", title = "Lead Time Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

From the graph, we can examine that lead_time varibale follows exponential distribution. <br/><br/> 

#### Question 9: Does lead time follow Central Limit Theorem?
```{r}
samples = 10000
sample.size1 = 10
xbar1 <- numeric(samples)
for(i in 1:samples){
  xbar1[i] <- mean(sample(hotel_bookings$lead_time, size = sample.size1, replace = TRUE))
}
xbar1 <- data.frame(xbar1)
p1 <- ggplot(xbar1, aes(x = xbar1)) + 
  geom_histogram(color = "orange", fill = "navyblue", lwd = 1, binwidth = 3) +
  theme_bw() +
  labs(x = "Lead Time Sample", y = "Frequency", title = "Sample Size = 10") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlim(0, 200)

sample.size2 = 50
xbar2 <- numeric(samples)
for(i in 1:samples){
  xbar2[i] <- mean(sample(hotel_bookings$lead_time, size = sample.size2, replace = TRUE))
}
xbar2 <- data.frame(xbar2)
p2 <- ggplot(xbar2, aes(x = xbar2)) + 
  geom_histogram(color = "orange", fill = "navyblue", lwd = 0.8, binwidth = 3) +
  theme_bw() +
  labs(x = "Lead Time Sample", y = "Frequency", title = "Sample Size = 50") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  xlim(0, 200)

sample.size3 = 100
xbar3 <- numeric(samples)
for(i in 1:samples){
  xbar3[i] <- mean(sample(hotel_bookings$lead_time, size = sample.size3, replace = TRUE))
}
xbar3 <- data.frame(xbar3)
p3 <- ggplot(xbar3, aes(x = xbar3)) + 
  geom_histogram(color = "orange", fill = "navyblue", lwd = 1, binwidth = 3) +
  theme_bw() +
  labs(x = "Lead Time Sample", y = "Frequency", title = "Sample Size = 100") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlim(0, 200)

sample.size4 = 500
xbar4 <- numeric(samples)
for(i in 1:samples){
  xbar4[i] <- mean(sample(hotel_bookings$lead_time, size = sample.size4, replace = TRUE))
}
xbar4 <- data.frame(xbar4)
p4 <- ggplot(xbar4, aes(x = xbar4)) + 
  geom_histogram(color = "orange", fill = "navyblue", lwd = 1, binwidth = 2) +
  theme_bw() +
  labs(x = "Lead Time Sample", y = "Frequency", title = "Sample Size = 500") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlim(0, 200)

grid.arrange(p1, p2, p3, p4)

# mean and standrad deviation values for each sample
sample.sizes <- c(10, 50, 100, 500)
means <- round(c(mean(xbar1$xbar1), mean(xbar2$xbar2), mean(xbar3$xbar3), mean(xbar4$xbar4)))
sds <- round(c(sd(xbar1$xbar1), sd(xbar2$xbar2), sd(xbar3$xbar3), sd(xbar4$xbar4)), 2)
for (i in 1:4){
  x <- sprintf("Sample Size = %g  Mean = %g  SD = %g", sample.sizes[i], means[i], sds[i])
  print(x)
}
```

By drawing samples from lead_time variable, we can see that the distribution of the sample means has the shape of a normal distribution for a large sample. The mean of the sample mean distribution is equal to the mean of the original lead_time data. The higher the sample, the smaller the standard deviation, the narrower the spread of the sample means.<br/><br/>


#### Question 10: Apply various sampling methods on lead_time data
```{r}
# simple random sampling
s <- srswr(1000, nrow(hotel_bookings))
rows <- (1:nrow(hotel_bookings))[s!=0]
rows <- rep(rows, s[s != 0])
sample1 <- hotel_bookings[rows, ]
sample1 <- data.frame(sample1$lead_time)
p1 <- ggplot(sample1, aes(x = sample1.lead_time)) +
        geom_histogram(color = "orange", fill = "navyblue", lwd = 0.8) +
        theme_bw() +
        labs(x = "Lead Time", y = "Frequency", title = "SRS (With Replacement)") +
        theme(plot.title = element_text(hjust = 0.5))

# simple random sampling without replacement
s <- srswor(1000, nrow(hotel_bookings))
sample2 <- hotel_bookings[s!=0, ]
sample2 <- data.frame(sample2$lead_time)
    p2 <- ggplot(sample2, aes(x = sample2.lead_time)) +
      geom_histogram(color = "orange", fill = "navyblue", lwd = 0.8) +
      theme_bw() +
      labs(x = "Lead Time", y = "Frequency", title = "SRS (Without Replacement)") +
      theme(plot.title = element_text(hjust = 0.5))

# Systemematic Sampling
N <- nrow(hotel_bookings)
n <- 1000
k <- ceiling(N/n)
r <- sample(k, 1)
s <- seq(r, by= k, length = n)
sample3 <- hotel_bookings[s, ]
sample3 <- data.frame(sample3$lead_time)
p3 <- ggplot(sample3, aes(x = sample3.lead_time)) + 
        geom_histogram(color = "orange", fill = "navyblue", lwd = 0.8) +
        theme_bw() +
        labs(x = "Lead Time", y = "Frequency", title = "Systematic Sampling") +
        theme(plot.title = element_text(hjust = 0.5))

# Unequal Probabilities
pik <- inclusionprobabilities(hotel_bookings$lead_time, 1000)
s <- UPsystematic(pik)
sample4 <- hotel_bookings[s != 0, ]
sample4 <- data.frame(sample4$lead_time)
p4 <- ggplot(sample4, aes(x = sample4.lead_time)) + 
        geom_histogram(color = "orange", fill = "navyblue", lwd = 0.8) +
        theme_bw() +
        labs(x = "Lead Time", y = "Frequency", title = "Unequal Probabilities") +
        theme(plot.title = element_text(hjust = 0.5))

# Stratified Sampling
freq <- table(hotel_bookings$hotel)
sizes <- 1000 * freq / sum(freq)
sample5 <- strata(hotel_bookings, stratanames = c("hotel"), size = sizes, method = "srswor",
                  description = FALSE)
sample5 <- getdata(hotel_bookings, sample5)
p5 <- ggplot(sample5, aes(x = lead_time)) + 
          geom_histogram(aes(fill = hotel)) +
          scale_fill_manual(name='Hotel Type', values =c('navyblue','orange')) +
          facet_wrap(~hotel) + 
          labs(x = "Lead Time", y = "Frequncy", title = "Stratified Sampling") +
            theme_bw() +
            theme(plot.title = element_text(hjust = 0.5))

p6 <- ggplot(hotel_bookings, aes(x = lead_time)) + 
        geom_histogram(aes(fill = hotel)) +
        scale_fill_manual(name='Hotel Type', values =c('navyblue','orange')) +
        facet_wrap(~hotel) + 
        labs(x = "Lead Time", y = "Frequncy", title = "Original Data") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5))
grid.arrange(p1, p2, p3, p4)
grid.arrange(p5, p6)

```

From the graph, we can see that samples from simple random sampling, systematic sampling and stratified sampling can represent the parent data due to similar shape and can be used for analysis. However, the data extracted by unequal probability method cannot represent the population and cannot be used for analysis.<br/><br/>

### PART4 Conclusion
1. Most customers who book hotels are adults, hotels may focus on designing rooms and promotions according to adults' preferences.  
2. The demand for parking space is very small, hotels may utilize some unused space to increase profits.  
3. The busiest period are late spring and early autumn.  
4. With booking amount increases, the cancel rate increases. With lead time increases, the cancel rate increases. Hotels may provide extra amount of rooms available for booking.  
5. City hotel has more customers compare to resort hotel does.  
6. Hotel may increase the price slightly in holodays and special days.  
